`r opts_chunk$set(cache=TRUE)`
# Chapter 2 - Single-parameter models
These notes explore various ideas discussed in Chapter 2 of Gelman et al *Bayesian Data Analysis*.

# Examining Binomial distribution
The following is just a refresher on the [binomial coefficient](http://en.wikipedia.org/wiki/Binomial_coefficient)

The following plots show the value for different values of $n$ and $k$.

```{r}
par(mfrow=c(3,3))
plot_choose <- function(n, ...) {
    plot(0:n, choose(n, 0:n), xlab='k', ylab='n choose k', ...)
}

for (X in c(2, 5, 8, 10, 20, 30, 50, 100, 200)) {
    plot_choose(X, main=paste('n =', X))
}
```

# Placenta privia example.

```{r}
german_data <- c(rep('female', 437), rep('male', 980 - 437))
table(german_data)
round(prop.table(table(german_data)), 3)

```


Examine beta distribution

The following plots $\text{Beta}(438, 544)$.

```{r}
Alpha <- 438
Beta <- 544
curve(dbeta(x, Alpha, Beta), 
      xlab=expression(theta),
      ylab=expression(paste("p(", theta, "|y)"))
      )
expected_beta <- Alpha / (Alpha + Beta)
var_beta <- (Alpha * Beta) /
    ((Alpha + Beta) ^2 * (Alpha + Beta + 1))
sd_beta <- sqrt(var_beta)
abline(v=expected_beta)

low_2.5_beta <- qbeta(.025, Alpha, Beta)
hi_2.5_beta <- qbeta(.975, Alpha, Beta)
abline(v=c(low_2.5_beta, hi_2.5_beta), lty=2)
```

Now let's compare beta to normal approximation

```{r}
low_2.5_normal <- qnorm(.025, expected_beta, sd_beta)
hi_2.5_normal <- qnorm(.975, expected_beta, sd_beta)

normal_beta <-cbind(`2.5`=c(low_2.5_normal, low_2.5_beta),
      `97.5`=c(hi_2.5_normal, hi_2.5_beta))
normal_beta <- rbind(normal_beta, apply(normal_beta, 2, diff))
row.names(normal_beta) <- c('normal', 'beta', 'diff')
round(normal_beta, 6)
```

* The normal approximation uses the mean and standard deviation of the Beta distribution in combination with the normal quantile function.
* This confirms what the authors state about the similarity between the exact betaq and the normal approximation of the 95% posterior interval.
* The values

Now let's simulate 1000 draws from the Beta distribution and calculate the 95% interval.

```{r}
set.seed(1234)
simulated_theta <- rbeta(1000, Alpha, Beta)
low_2.5_sim_beta <- sort(simulated_theta)[25]
hi_97.5_sim_beta <-sort(simulated_theta)[976]

rbind(normal_beta, sim_beta=c(low_2.5_sim_beta, hi_97.5_sim_beta))

median(simulated_theta)

c(simulated=mean(simulated_theta), expected=expected_beta)

c(simulated=sd(simulated_theta), sd=sd_beta)
```

* It might be interesting to compare results for smaller samples or for values of $\theta$ closer to zero or one.

# Exploring table 2.1
```{r}
what_alpha_beta <- function(a_over_a_plus_b, a_plus_b) {
    # determines alpha and beta for a beta function where
    # alpha / (alpha + beta) and
    # alpha + beta
    # are known
    
    Alpha <- a_over_a_plus_b / a_plus_b
    Beta <- a_plus_b - Alpha
    c(Alpha=Alpha, Beta=Beta)
}

table_2.1 <- cbind(one=c(0.5, rep(.485, 6)),
        two=c(2,2,5,10,20,100,200))

table_2.1 <- cbind(table_2.1, 
                   t(apply(table_2.1, 1, 
                           function(X) what_alpha_beta(X[1], X[2]))))

table_2.1
# I haven't completed this analysis
```

# Piecewise linear density
What are some examples of piecewise linear densities?

Here's one strategy for constructing a piecewise density using increments and decrements and a cumulative total. However, it's normalised for discrete data, which is a limitation. 

```{r}
Data <- data.frame(theta=seq(0, 1, .01), 
               raw_p=cumsum(c(1, rep(0, 40), rep(1, 10), rep(-1, 10), rep(0, 40))))

Data$normalised_p <- Data$raw_p / sum(Data$raw_p) 
plot(Data$theta, Data$normalised_p, type='l')
```

# Asthma death example
```{r}
Alpha <- 3
Beta <- 5

curve(dgamma(x, Alpha, Beta), from=0, to=4)
abline(v=Alpha /Beta)
abline(v=(Alpha  -1 ) / Beta)
text(2,1, paste('E(X)=',Alpha /Beta))
text(2,.9, paste('var(X)=', Alpha /Beta^2))
text(2,.8, paste('mode(X)=', (Alpha  -1 ) / Beta))

par(mfrow=c(3,3))
curve(gamma(x), from=0.5, to=4)
curve(gamma(x) * Beta ^ x, from=0.5, to=4)
curve(1/ (gamma(x) * Beta ^ x), from=0.5, to=4)

curve(Beta^x, from=0.5, 4)
curve(exp(-Beta/x), from=0.5, to=4)
curve(exp(-Beta/x), from=0.5, to=400)

# Form used in Gelman
curve(exp(-Beta * x), from=0.5, to=4)
curve(exp(-Beta *x), from=0.5, to=400)
```

Note that there are different parameterisations of the Gamma distribution.

* Some people use $e^{-\theta/\beta}$ (i.e., a scale parameter) whereas Gelman et al uses $e^{-\beta\theta}$ (i.e., a rate parameter).
* This changes $E(X) = \alpha\beta$ to $E(X) = \frac{\alpha}{\beta}$.
* The `dgamma` function defaults to a rate parameter, but also has an optional `scale` parameter that can be used.


```{r}
# sample from posterior
x <- rgamma(1000, 6, 7)
hist(x, 20, freq=FALSE)
abline(v=mean(x))
lines(density(x))

```

# Exponential distribution

```{r}
par(mfrow=c(2,2))
Betas <- 1:4
for (Beta in Betas) {
curve(dexp(x, Beta), 0, 10)
text(4, 1, paste('E(X)=', Beta))
text(4, 0.6, paste('var(X)=', Beta^2))
text(4, 0.4, paste('mode(X)=', 0))
abline(v=1)
}
```

# Example 2.8
```{r}
curve(dgamma(x, 20, 4.3E+5), from=0, to=.0001)
set.seed(1234)
thetas <- rgamma(500, 20, 4.3E+5)
ys <- rpois(thetas, thetas*1.0E+4)
mean(ys/1.0E+4)
table(ys)
prop.table(table(ys))

thetas <- rgamma(500, 20, 4.3E+5)
ys <- rpois(thetas, thetas*1.0E+6)
mean(ys/1.0E+6)
table(ys)
prop.table(table(ys))
```

# Inverse chi-square distribution
```{r}
par(mfrow=c(3,3))
curve(exp(x), 0, 10)
curve(exp(-x), 0, 10)
curve(exp(1/x), 0, 0.01)
curve(exp(-1/x), 0, 10)
curve(-1/x, 0, 10)
curve(-1/(2*x), 0, 10)
curve(exp(-1/(2*x)), 0, 10)
```

* $\exp(0)=1$
* $\lim_{x \to -\infty} \exp(x) = 0$
* $\lim_{x \to \infty} -1/x = 0$
* $\lim_{x \to \infty} \exp(-1/x) = 1$


```{r}
par(mfrow=c(3,3))
curve(x^(-(1/2 + 1)), 0, 10)
```

```{r}
par(mfrow=c(3,3))
plot_inv_chisquare_kernal <- function(Nu) {
    curve(x^(-Nu/2-1)* exp(-1/(2*x)), 0, 5)    
}

plot_inv_chisquare_kernal(3)
plot_inv_chisquare_kernal(2)
plot_inv_chisquare_kernal(1)
plot_inv_chisquare_kernal(1)

curve(exp(x^(10/2-1)* (-1/(10*x))), 0, 10)
```

```{r}
integrate(function(x) 1/x^2, 1, Inf)
integrate(function(x) 1/x^2, 0.0001, Inf)
integrate(function(x) 1/x^2, 0, Inf)
```


